Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-12-13 ms.tex
Rewording the steady-state discussion section, changing ordering of sentences

\subsection{Steady State Source}
A primary goal of this work is to present a solution for resonant scattering near the line-center frequency $\nu_0$ in a uniform sphere. We have generalized a spherically symmetric solution derived by \cite{2006ApJ...649...14D} (called $H_0$ here) to allow a monochromatic source of photons with frequencies away from line center. We have included a correction to their incorrect $J=0$ surface boundary condition that allows the proper condition $J=\sqrt{3}H$ to be satisfied at the surface of the sphere. This introduces a term $J_{\rm bc}$ which is solved using a continuous Fourier expansion in frequency. The continuous expansions are discretized and the Fourier coefficients solved for. The resulting flux correction, $H_{\rm bc}$, scales as $H_0(a\tau_0)^{-1/3}$. Thus, for large $\tau_0$, only a small correction to $H_0$ is needed, while larger errors are present in calculations performed at lower $\tau_0$. Since the Laplacian form of the differential equation is only correct when the line profile is $\phi \approx a/(\pi x^2)$, our solutions do not accurately model the Doppler core of the \lya line. Because the peak of the spectral energy distribution of escaping photons is $x_{\rm peak} = (a\tau_0)^{1/3}$, calculations performed at low $\tau_0$ are inaccurate due to the close proximity of the spectral peak and the Doppler core of the line.

By comparison with Monte Carlo simulations, we have shown that the enforcement of the correct frequency-dependent boundary conditions improves the accuracy of these analytic solutions provided a sufficiently large $\tau_0$. Specifically, this solution shows improvement over previous solutions that utilized the $J=0$ surface boundary condition presented in \cite{1973MNRAS.162...43H} and \cite{1990ApJ...350..216N}. Several papers have previously compared these analytic models to Monte Carlo and seen discrepancies on the order of this correction. For example, in the top-left panel of Figure 1 from \cite{2006ApJ...649...14D}, the \lya spectrum emergent from a sphere of uniform optical depth is shown for $\tau_0=10^5, 10^6,$ and $10^7$ at a temperature of $T=10$ K, corresponding to $a=1.5 \times 10^{-2}$. The dotted line showing their theoretically-derived spectrum (called $H_0$ here) displays an excess at the peak of at least 5-10 percent as compared with the Monte Carlo for $\tau_0=10^5$ and $10^6$. Though our calculations were performed with a higher gas temperature of $T=10^4$ K, we show in Figure \ref{fig:sol_mc_tau} that the excess present in $H_0$ is corrected by our treatment of the boundary condition at $\tau_0 = 10^7$. In \citet{2015MNRAS.449.4336S}, the peak excess in the \lya spectrum is particularly noticeable for line center optical depths of $\tau_0=10^6$ and $10^7$ in the top panel of their Figure 5, which used slab geometry and a gas temperature of $T=10^4$ K. Again, the error in their solution is of order 5-10 percent. $H_{\rm bc}$ is positive in the line wing and negative at the peak of the spectrum, which matches with the discrepancies noted in these solutions. They are too large at the spectral peaks and too small further out in the wing, and the error scales approximately as $(a\tau_0)^{-1/3}$.

>>>>

\subsection{Steady State Source}
A primary goal of this work is to present a solution for resonant scattering of photons near the line-center frequency $\nu_0$ in a uniform sphere. We have generalized a spherically symmetric solution derived by \cite{2006ApJ...649...14D} (called $H_0$ here) to allow a monochromatic source of photons with frequencies away from line center. We introduce a new term to this solution, $J_{\rm bc}$, which allows the boundary condition $J=\sqrt{3}H$ to be satisfied at the surface of the sphere. This is solved using a continuous Fourier expansion in frequency. The integrals are discretized and the Fourier coefficients solved for numerically. The resulting flux correction, $H_{\rm bc}$, scales as $H_0(a\tau_0)^{-1/3}$. Thus, for large $a\tau_0$, only a small correction to $H_0$ is needed, while larger errors are present in calculations performed at lower $a\tau_0$. Since the Laplacian form of the differential equation is only correct when the line profile is $\phi \approx a/(\pi x^2 \Delta)$, our solutions do not accurately model the Doppler core of the \lya line. Because the peak of the spectral energy distribution of escaping photons is $x_{\rm peak} {\sim} (a\tau_0)^{1/3}$, calculations performed at small $a\tau_0$ are inaccurate due to the close proximity of the spectral peak and the Doppler core of the line.

By comparison with Monte Carlo simulations, we have shown that the enforcement of the correct frequency-dependent boundary conditions improves the accuracy of these analytic solutions provided a sufficiently large $a\tau_0$. Specifically, this solution shows improvement over previous solutions that utilized a $J=0$ surface boundary condition presented in \cite{1973MNRAS.162...43H}, \cite{1990ApJ...350..216N}, and \cite{2006ApJ...649...14D}. Several papers have previously compared these analytic models to Monte Carlo and seen discrepancies on the order of this correction. For example, in the top-left panel of Figure 1 from \cite{2006ApJ...649...14D}, the \lya spectrum emergent from a sphere of uniform optical depth is shown for $\tau_0=10^5, 10^6,$ and $10^7$ at a temperature of $T=10$ K, corresponding to $a=1.5 \times 10^{-2}$. The dotted line showing their theoretically-derived spectrum ($H_0$) displays an excess at the peak of at least 5-10 percent as compared with the Monte Carlo for $\tau_0=10^5$ and $10^6$. Another example is in \citet{2015MNRAS.449.4336S}, where the peak excess in the \lya spectrum is particularly noticeable for line center optical depths of $\tau_0=10^6$ and $10^7$ in the top panel of their Figure 5, which used slab geometry and a gas temperature of $T=10^4$ K. Again, the error in their solution is of order 5-10 percent. Both of these solutions are too large at the spectral peaks and too small further out in the wing, and the error scales approximately as $(a\tau_0)^{-1/3}$. We show in our Figure \ref{fig:sol_mc_tau} that the error present in $H_0$ is corrected by our treatment of the boundary condition at $\tau_0 = 10^7$ for T$=10^4$ K, corresponding to $a=4.72 \times 10^{-4}$. We note that our corrective term $H_{\rm bc}$ is positive in the line wing and negative at the peak of the spectrum, which matches with the discrepancies noted in the aforementioned solutions.
